On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-08-26 paper.tex
Got rid of "a" in the square grid with real-valued lengths

Knowing a minimal sufficient set of $(i,j)$ pairs facilitates the solution of both the direct an inverse problems, because it reduces the number of pairs that need to be tried. In order to derive a general method to build a minimal sufficient set, it is insightful to consider two specific cases first, as will be presented in the following. It is also helpful to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.\ finding the maximum number of tiles visited by segments of length $\len$ \emph{or less}.

@@

\subsection{Square grid with real-valued lengths}
\label{part: max: square grid, real lengths}

Particularizing the results in \S\ref{part: max: arbitrary grid, real lengths} to a square grid, $a=b$, obviously yields simpler formulas.

\begin{corollary}
\label{cor: funt, sq, form}
For a square grid with $a \in \mathbb R^+$, and for $\len \in \mathbb R^+$,
\begin{equation}
\label{eq: cor: funt, sq, form}
\funt(\len) = \isoli+\jsoli-1
\end{equation}
with
\begin{align}
\label{eq: cor: funt, sq, form; i}
\isoli &= \left\lceil \frac{1 + \Re \sqrt{2\len^2/a^2-1}}{2} \right\rceil + 1, \\
\label{eq: cor: funt, sq, form; j}
\jsoli &= \left\lceil \sqrt{\len^2/a^2-(\isoli-2)^2} \right\rceil + 1,
\end{align}
\end{corollary}

\begin{corollary}
\label{theo: funl, sq, form}
For a square grid with $a \in \mathbb R^+$, and for $\tiles \in \mathbb N$,
\begin{equation}
\label{eq: theo: funt, sq, form; funl}\\
\frac{\funl(\tiles)}{a} = \begin{cases}
\displaystyle
0 & \text{for } \tiles =1, 2 \\[1.4mm]
\displaystyle
\frac{\tiles-3}{\sqrt{2}} & \text{for } \tiles \text{ odd, } \tiles \geq 3 \\[4.5mm]
\displaystyle
\sqrt{\frac{(\tiles-3)^2+1} {2}} & \text{for } \tiles \text{ even, } \tiles \geq 3.
\end{cases}
\end{equation}
\end{corollary}

@@

\begin{theorem}
\label{theo: funt, sq, form}
For a square grid with $a \in \mathbb R^+$, and for $\len \in \mathbb R^+$,
\begin{equation}
\label{eq: theo: funt, sq, form; funt}
\funt(\len) = \isoli+\jsoli-1
\end{equation}
with
\begin{align}
\label{eq: theo: funt, sq, form; i}
\isoli &= \left\lceil \frac{\len}{a \sqrt{2}} \right\rceil + 1, \\
\label{eq: theo: funt, sq, form; j}
\jsoli &= \left\lceil \sqrt{\len^2/a^2-(\isoli-2)^2} \right\rceil + 1.
\end{align}
\end{theorem}

\begin{proof}
For $a=b$ the set $\mss$ from Proposition~\ref{prop: min suff set} consists of points of the form $(i,i)$ and $(i+1,i)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}. By symmetry, replacing each point $(i+1,i)$ by $(i,i+1)$ gives a set $\mss'$ which is also minimal sufficient. For this new set, the lower bounding line \eqref{eq: lower bound, line} can be replaced by the simpler $j=i$. The same approach used in the proof of Theorem~\ref{theo: funt, form} can be used, but using this line. Thus $(\isolr, \jsolr)$ is obtained from
\begin{align}
(\isolr-2)^2 a^2 + (\jsolr-2)^2 a^2 &= \len^2, \\
\jsolr &= \isoli,
\end{align}
which gives
\begin{equation}
\label{eq: theo: funt, sq, form; proof 1}
\isolr = \frac{\len}{a \sqrt{2}} + 2.
\end{equation}
The pair $(\isoli,\jsoli)$ resulting from \eqref{eq: theo: funt, sq, form; proof 1} is in $\mss'$, and is given in \eqref{eq: theo: funt, sq, form; i} and \eqref{eq: theo: funt, sq, form; j}.
\end{proof}

From Theorem~\ref{theo: funt, sq, form} it is easily seen that odd values of $\funt(\len)$ correspond to $\isoli = \jsoli$, whereas even values are achieved with $\jsoli = \isoli+1$. Given $\tiles \geq 3$ with $\tiles$ odd, $\funt(\len) = \tiles$ if and only
\begin{equation}
\label{eq: square, odd max tiles, lengths}
\frac \len a \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{ \sqrt{(\tiles-1)^2+(\tiles-3)^2}}{2} \right].
\end{equation}
For $\tiles \geq 4$, $\tiles$ even, $\funt(\len) = \tiles$ if and only if
\begin{equation}
\label{eq: square, even max tiles, lengths}
\frac \len a \in \left(\frac{ \sqrt{(\tiles-2)^2+(\tiles-4)^2}}{2}, \frac{\tiles-2}{\sqrt{2}} \right].
\end{equation}

@@

A natural variation of the two problems introduced in \S\ref{part: intro} is to consider $a=b=1$, with the additional restriction that the segment length can only be a positive integer (equivalently, a square grid is considered with step $a$ and the segment length can only be an integer multiple of $a$).

>>>>

Knowing a minimal sufficient set of $(i,j)$ pairs facilitates the solution of both the direct an inverse problems, because it reduces the number of pairs that need to be tried. In order to derive a general method to build a minimal sufficient set, it is insightful to consider two specific cases first, as will be presented in the following. It is also helpful to use the equivalent formulation of the direct problem referred to in \S\ref{part: intro}, i.e.~finding the maximum number of tiles visited by segments of length $\len$ \emph{or less}.

@@

\subsection{Unit square grid with real-valued lengths}
\label{part: max: square grid, real lengths}

A square grid has $a=b$. If the segment lengths are real-valued it can be further assumed that $a=1$. For $a \neq 1$ the results to be obtained apply with $\len$ replaced by $\len/a$.

Particularizing the results in \S\ref{part: max: arbitrary grid, real lengths} to $a=b=1$ obviously yields simpler formulas.

\begin{corollary}
\label{cor: funt, sq, form}
For a unit square grid with $\len \in \mathbb R^+$,
\begin{equation}
\label{eq: cor: funt, sq, form}
\funt(\len) = \isoli+\jsoli-1
\end{equation}
with
\begin{align}
\label{eq: cor: funt, sq, form; i}
\isoli &= \left\lceil \frac{1 + \Re \sqrt{2\len^2-1}}{2} \right\rceil + 1, \\
\label{eq: cor: funt, sq, form; j}
\jsoli &= \left\lceil \sqrt{\len^2-(\isoli-2)^2} \right\rceil + 1,
\end{align}
\end{corollary}

\begin{corollary}
\label{theo: funl, sq, form}
For a unit square grid, and for $\tiles \in \mathbb N$,
\begin{equation}
\label{eq: theo: funt, sq, form; funl}\\
\funl(\tiles) = \begin{cases}
\displaystyle
0 & \text{for } \tiles =1, 2 \\[1.4mm]
\displaystyle
\frac{\tiles-3}{\sqrt{2}} & \text{for } \tiles \text{ odd, } \tiles \geq 3 \\[4.5mm]
\displaystyle
\sqrt{\frac{(\tiles-3)^2+1} {2}} & \text{for } \tiles \text{ even, } \tiles \geq 4.
\end{cases}
\end{equation}
\end{corollary}

@@

\begin{theorem}
\label{theo: funt, sq, form}
For a unit square grid with $\len \in \mathbb R^+$,
\begin{equation}
\label{eq: theo: funt, sq, form; funt}
\funt(\len) = \isoli+\jsoli-1
\end{equation}
with
\begin{align}
\label{eq: theo: funt, sq, form; i}
\isoli &= \left\lceil \frac{\len}{\sqrt{2}} \right\rceil + 1, \\
\label{eq: theo: funt, sq, form; j}
\jsoli &= \left\lceil \sqrt{\len^2-(\isoli-2)^2} \right\rceil + 1.
\end{align}
\end{theorem}

\begin{proof}
For $a=b=1$ the set $\mss$ from Proposition~\ref{prop: min suff set} consists of points of the form $(i,i)$ and $(i+1,i)$, as is easily seen from Proposition~\ref{prop: min suff set, form}, and as illustrated in Figure~\ref{fig: pairs_bounds_1}. By symmetry, replacing each point $(i+1,i)$ by $(i,i+1)$ gives a set $\mss'$ which is also minimal sufficient. For this new set, the lower bounding line \eqref{eq: lower bound, line} can be replaced by the simpler $j=i$. The same approach used in the proof of Theorem~\ref{theo: funt, form} can be followed, but using this line. Thus $(\isolr, \jsolr)$ is obtained from
\begin{align}
(\isolr-2)^2 + (\jsolr-2)^2  &= \len^2, \\
\jsolr &= \isoli,
\end{align}
which gives
\begin{equation}
\label{eq: theo: funt, sq, form; proof 1}
\isolr = \len/\sqrt{2} + 2.
\end{equation}
The pair $(\isoli,\jsoli)$ resulting from \eqref{eq: theo: funt, sq, form; proof 1} is in $\mss'$, and is given in \eqref{eq: theo: funt, sq, form; i} and \eqref{eq: theo: funt, sq, form; j}.
\end{proof}

From Theorem~\ref{theo: funt, sq, form} it is easily seen that odd values of $\funt(\len)$ correspond to $\isoli = \jsoli$, whereas even values are achieved with $\jsoli = \isoli+1$. Given $\tiles \geq 3$ with $\tiles$ odd, $\funt(\len) = \tiles$ if and only
\begin{equation}
\label{eq: square, odd max tiles, lengths}
\len \in \left( \frac{\tiles-3}{\sqrt{2}}, \frac{ \sqrt{(\tiles-1)^2+(\tiles-3)^2}}{2} \right].
\end{equation}
For $\tiles \geq 4$, $\tiles$ even, $\funt(\len) = \tiles$ if and only if
\begin{equation}
\label{eq: square, even max tiles, lengths}
\len \in \left(\frac{ \sqrt{(\tiles-2)^2+(\tiles-4)^2}}{2}, \frac{\tiles-2}{\sqrt{2}} \right].
\end{equation}

@@

A natural variation of the first two problems introduced in \S\ref{part: intro} is to consider $a=b=1$, with the additional restriction that the segment length can only be a positive integer (equivalently, a square grid is considered with step $a$ and the segment length can only be an integer multiple of $a$).
